### 2024-05-19 chapters/data-analysis.tex
chore(sections): rename to data-*

\chapter{Anàlisi i visualització de les dades}\label{ch:log-analysis}