### 2024-05-21 appendices/software-installation.tex
chore(syle): delete number from toc in specific chapters

\chapter{Instal·lació de software}\label{ch:software-installation}