### 2024-05-21 chapters/introduction.tex
chore(style): use EPSEVG final thesis template (#6)
* feat(titlepage): add epsevg template

* feat(titlepage-epsevg): update title name

* feat(titlepage-epsevg): update title name

* feat(titlepage): add thesis evaluation page

* feat(abstract): add ca-es-en abstracts

* feat(acks): add aknowledgments page

* feat(main): update main files

* fix(typo)

* feat(thesis-evaluation): increase box width

* feat(style): fix frontmatter style

* chore(syle): delete number from toc in specific chapters

* feat(glossary): add glossary

* feat(metadata): add new source command

\chapter*{Introducció}\label{ch:introduction}
\addcontentsline{toc}{chapter}{Introducció}