### 2024-05-22 chapters/data-storing.tex
feat(chapters): add log-storing chapter (#12)

\chapter{Emmagatzematge de les dades}\label{ch:log-storing}