### 2024-05-22 chapters/data-processing.tex
feat(figures): add log-processing-workflow

\chapter{Processament de les dades}\label{ch:log-processing}

\input{sections/metadata}
\clearpage
\input{sections/logs}