### 2024-05-22 sections/logs.tex

\section{Logs}\label{sec:logs}

\begin{figure}[htbp]
    \centerline{\includegraphics[width=1.3\textwidth]{figures/log-processing-workflow}}
    \captionsetup{justification=centering}
    \caption{Implementació del processament dels \textit{logs}.}\label{fig:log-processing}
    \source{Elaboració pròpia.}
\end{figure}