### 2024-05-22 sections/metadata.tex

\section{Metadades}\label{sec:metadata}

\begin{figure}[htbp]
    \centerline{\includegraphics[width=1.2\textwidth]{figures/metadata-processing}}
    \captionsetup{justification=centering}
    \caption{Disseny tècnic del processament de les metadades.}\label{fig:log-analysis}
    \source{Elaboració pròpia.}
\end{figure}