### 2024-05-28 chapters/data-processing.tex
Update subsection file name

\chapter{Processament de les dades}\label{ch:log-processing}

\input{sections/logs-processing}
\clearpage
\input{sections/metadata-processing}