### 2024-05-28 chapters/data-storing.tex
Add data processing section

\chapter{Emmagatzematge de les dades}\label{ch:log-storing}

\input{sections/logs-storing}
\clearpage
\input{sections/metadata-storing}