### 2024-05-28 tables/metadata-push/metadata-mongodb.tex

\begin{table}[h!]
    \centering
    \begin{adjustbox}{width=0.9\textwidth, center}
        \begin{tabular}{|l|r|r|}
            \toprule
            \begin{tabular}[c]{@{}l@{}}   \textbf{\texttt{/mnt/working/metadata}} → \\ \textbf{MongoDB}             \end{tabular}
            & \begin{tabular}[c]{@{}l@{}} Nº total de \\ metadades                                \end{tabular}
            & \begin{tabular}[c]{@{}l@{}} Duració de \\ l’abocament (segons)                     \end{tabular}   \\
            \midrule
            \textbf{Total}  & 242.555   &   38,2050559520721 \\
            \bottomrule
        \end{tabular}
    \end{adjustbox}
    \caption{Resultat de l'abocament de les metadades a \textbf{MongoDB}}
    \label{tab:metadata-push-mongodb}
\end{table}