### 2024-05-28 tables/metadata-push/metadata-filesystem.tex
Update metadata storing section (#20)

\begin{table}[h!]
    \centering
    \begin{adjustbox}{width=0.9\textwidth, center}
        \begin{tabular}{|l|r|r|}
            \toprule
            \begin{tabular}[c]{@{}l@{}}   \textbf{Servidor OAI} → \\ \textbf{\texttt{/mnt/working/metadata}}        \end{tabular}
            & \begin{tabular}[c]{@{}l@{}} \textbf{Nº total de} \\ \textbf{metadades}                                \end{tabular}
            & \begin{tabular}[c]{@{}l@{}} \textbf{Duració de} \\ \textbf{l’abocament (minuts)}                      \end{tabular}   \\
            \midrule
            \textbf{Total}  & 242.555   &   20,95360236565 \\
            \bottomrule
        \end{tabular}
    \end{adjustbox}
    \caption{Resultat de l'abocament de les metadades a \texttt{/mnt/working/metadata}}
    \label{tab:metadata-push-filesystem}
\end{table}