### 2024-05-31 chapters/abstract-en.tex
delete lipsum and methodology

\chapter*{Abstract}\label{ch:abstract-en}

\noindent \\
\section*{Keywords}\label{sec:keywords-en}
%\begin{multicols}{2}
%    \begin{itemize}
%        \item Keyword
%        \item Keyword
%        \item Keyword
%        \item Keyword
%    \end{itemize}
%    \columnbreak
%    \begin{itemize}
%        \item Keyword
%        \item Keyword
%        \item Keyword
%        \item Keyword
%    \end{itemize}
%\end{multicols}