### 2024-05-31 chapters/abstract-ca.tex
feat(thesis): fix typos (#27)
* fix typos

* remove typos

* fix typos

* delete lipsum and methodology

* update bibliography dates

* fix typos

\chapter*{Resum}\label{ch:abstract-ca}

\noindent \\
\section*{Paraules clau}\label{sec:keywords-ca}
%\begin{multicols}{2}
%    \begin{itemize}
%        \item Paraula
%        \item Paraula
%        \item Paraula
%        \item Paraula
%    \end{itemize}
%    \columnbreak
%    \begin{itemize}
%        \item Paraula
%        \item Paraula
%        \item Paraula
%        \item Paraula
%    \end{itemize}
%\end{multicols}