### 2024-05-31 chapters/abstract-es.tex

\chapter*{Resumen}\label{ch:abstract-es}

\noindent \\
\section*{Palabras clave}\label{sec:keywords-es}
%\begin{multicols}{2}
%    \begin{itemize}
%        \item Palabra
%        \item Palabra
%        \item Palabra
%        \item Palabra
%    \end{itemize}
%    \columnbreak
%    \begin{itemize}
%        \item Palabra
%        \item Palabra
%        \item Palabra
%        \item Palabra
%    \end{itemize}
%\end{multicols}